Body Mass Index and Pathway Dysregulation in Cancer

### 2016-10-22 presentation/beamer_theme/ctcrSep2016.tex
tmp commit

\title[BMI and Cancer]{Body Mass Index and Pathway Dysregulation in Cancer}
\subtitle{CTCR Talk}
\author[Riku Takei]{Riku Takei (MSc)\\Supervisor: Assoc. Prof. Mik Black}
\institute{Department of Biochemistry, University of Otago}

@@

\begin{frame}
	\frametitle{Methods}
	Downloaded RNA-seq data from 8 different cancer types from ICGC
	\vskip2ex
	Microarray data from Creighton's paper
	\vskip2ex
	Microarray data from Fuentes-Mattei's paper
	\vskip2ex
	Microarray data from Cris Print
	\vskip2ex
	\includegraphics[width=0.4\textwidth]{./images/tcga}
	\hfill
	\includegraphics[width=0.4\textwidth]{./images/icgc}
\end{frame}

@@

\begin{frame}{Creighton \textit{et al.} data revisited}
	Gene expression analysis was carried out on the Creighton \textit{et al.} data:
	\begin{itemize}
		\item Original Creighton data
		\item Caucasian-only Creighton data
	\end{itemize}
	\vskip2ex
	Gene expression analysis was repeated with the residual version of the above data
	\vskip2ex
	Each ``obesity-associated genes'' were checked for any overlaps with the original Creighton metagene
\end{frame}

@@

\begin{frame}{Metagene Analysis}
	All of the gene probe sets I have found (8 in total) were used to create metagenes as before
	\begin{itemize}
		\item Checked in Creighton \textit{et al.} data
		\item Transferred into other cancer data sets
	\end{itemize}
\end{frame}

@@

\begin{frame}{Relevance of BMI in Cris' data}
	\begin{center}
		\includegraphics[width=0.60\paperwidth]{./images/tmp/mikres}
	\end{center}
\end{frame}

@@

\begin{frame}{Gatza \textit{et al.} (2010)}
	Used 18 pathways to define sub-class of breast tumours, based on the expression of the pathway signatures
	\begin{itemize}
		\item AKT, ER, PR, HER2, PI3K, p53, MYC, RAS, etc...
	\end{itemize}
	\vskip2ex
	Each pathway signature was obtained by overexpressing a single gene in an appropriate cell line, and then differentially expressed genes were noted down
	\begin{itemize}
			\item Use these pathways to see what the obesity-associated genes are most similar to
	\end{itemize}
\end{frame}

@@

\begin{frame}{Predicting obesity metagene with pathway metagene}
	Make a linear model (in Cris' data) with sample BMI and/or pathway metagenes and use this to predict Creighton's obesity metagene
	\vskip2ex
	\begin{itemize}
		\item If any of the pathway metagenes are significant in the model, the pathway is a good predictor of the obesity metagene
	\end{itemize}
	\vskip2ex
	Then use this model to predict the obesity metagene in Creighton data
	\begin{itemize}
		\item Compare the predicted obesity metagene with the SVD-generated obesity metagene
	\end{itemize}
\end{frame}

\begin{frame}{Results}
	Sample BMI was not significant  in the model
	\vskip2ex
	Only pathway that was a good predictor of obesity metagene was PR pathway metagene
\end{frame}

@@

\begin{frame}{Overall Summary}
	\textbf{Creighton/FM metagene analysis}
	\begin{itemize}
		\item Both Creighton and FM metagenes correlated with the overall gene expression in any cancer data set
		\item The metagene score does not correlate with the BMI status of the samples
	\end{itemize}
	\vskip2ex
	\textbf{Gene expression and metagene analysis}
	\begin{itemize}
		\item Again, all of the gene sets found correlated with the overall gene expression in any data set, but none of them correlated with the BMI status
	\end{itemize}
\end{frame}

\begin{frame}{Overall Summary}
	\textbf{Gatza pathway metagene analysis}
	\begin{itemize}
		\item Sample BMI was not a good predictor of the obesity metagene, but PR pathway metagene was
		\item However, the model was not great at predicting the obesity metagene
			\begin{itemize}
				\item Other mechanism contributing to obesity metagene?
			\end{itemize}
	\end{itemize}
\end{frame}

>>>>

\title[BMI and Cancer]{Body Mass Index and Pathway Dysregulation in Cancer}
\subtitle{CGL Talk}
\author[Riku Takei]{Riku Takei (MSc)\\Supervisor: Assoc. Prof. Mik Black}
\institute{Department of Biochemistry, University of Otago}

@@

\begin{frame}
	\frametitle{Methods}
	Downloaded RNA-seq data from 8 different cancer types from ICGC
	\vskip2ex
	Microarray data from:
	\begin{itemize}
		\item Creighton's paper
		\item Fuentes-Mattei's paper
		\item Cris Print
	\end{itemize}
	\vskip2ex
	\includegraphics[width=0.4\textwidth]{./images/tcga}
	\hfill
	\includegraphics[width=0.4\textwidth]{./images/icgc}
\end{frame}

@@

\begin{frame}{Creighton \textit{et al.} data revisited}
	Gene expression analysis was carried out on:
	\begin{itemize}
		\item Original Creighton data
		\item Creighton data (Caucasian samples only)
	\end{itemize}
	\vskip2ex
	Gene expression analysis was repeated with the residual version of the above data
	\vskip2ex
	Each ``obesity-associated genes'' were checked for any overlaps with the original Creighton metagene
\end{frame}

@@

\begin{frame}{Metagene Analysis}
	All of the gene probe sets I have found (8 in total) were used to create metagenes as before
	\begin{itemize}
		\item Checked in Creighton \textit{et al.} data
		\item Transferred into other cancer data sets using transformation matrix
	\end{itemize}
\end{frame}

@@

\begin{frame}{So \ldots BMI is not related to cancer?}
	\begin{center}
		\includegraphics[width=0.55\paperwidth]{./images/tmp/mikres}
	\end{center}
\end{frame}

@@

\begin{frame}{Gatza \textit{et al.} (2010)}
	Used 18 pathways to define sub-classes of breast tumours, based on the expression of the pathway signatures
	\begin{itemize}
		\item AKT, ER, PR, HER2, PI3K, p53, MYC, RAS, etc...
	\end{itemize}
	\vskip2ex
	Each pathway signature was obtained by overexpressing a single gene in a model cell line, and then differentially expressed genes were noted down
	\begin{itemize}
			\item Use these `pathway genes' to see what the obesity-associated genes are most similar to
	\end{itemize}
\end{frame}

@@

\begin{frame}{\normalsize Predicting obesity metagene with pathway metagene}
	Make a linear model (in Cris' data) with sample BMI and/or pathway metagenes and use this to predict Creighton's obesity metagene
	\begin{itemize}
		\item If any of the pathway metagenes are significant in the model, the pathway is a good predictor of the obesity metagene
	\end{itemize}
	\vskip2ex
	Then use this model to predict the obesity metagene in Creighton data
	\begin{itemize}
		\item Compare the predicted obesity metagene with the SVD-generated obesity metagene
	\end{itemize}
\end{frame}

\begin{frame}{Results}
	Made 4 models:
	\begin{itemize}
		\item BMI-only
		\item BMI and pathway metagenes
		\item Pathways-only
		\item PR pathway only
	\end{itemize}
	\vskip2ex
	Sample BMI was not significant  in the model
	\vskip2ex
	Only pathway that was a good predictor of obesity metagene was PR pathway metagene
\end{frame}

@@

\begin{frame}{Overall Summary}
	\textbf{Creighton/FM metagene analysis}
	\begin{itemize}
		\item Both Creighton and FM metagenes correlated with the overall gene expression in any cancer data set
		\item The metagene score does not correlate with the BMI status of the samples
	\end{itemize}
	\vskip2ex
	\textbf{Gene expression and metagene analysis}
	\begin{itemize}
		\item Again, all of the gene sets correlated with the overall gene expression in any data set, but none of them correlated with the BMI status
	\end{itemize}
\end{frame}

\begin{frame}{Overall Summary}
	\textbf{Gatza pathway metagene analysis}
	\begin{itemize}
		\item Sample BMI was not a good predictor of the obesity metagene, but PR pathway metagene was
		\item None of the models were great at predicting the obesity metagene
			\begin{itemize}
				\item Other mechanism contributing to obesity metagene?
			\end{itemize}
	\end{itemize}
\end{frame}
